---
title: "Deep dive into deep-sea coral and sponge research project outcomes"
author: "Sydney Lewis"
date: "2022-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in Google Sheet

install.packages('googlesheets4')

```{r cars}
#Load the required library 
library(googlesheets4)

#Read google sheets data into R
dsc <- read_sheet("https://docs.google.com/spreadsheets/d/1ndur-uZ5D-YHnGYjh913t4OM76X-GrWuS1-y2XazLSs/edit?usp=sharing")
```

Set up subset of dataset to work with 
```{r pressure, echo=FALSE}
library(dplyr)

#Remove columns with character values/ notes & links 
dsc1<- subset(dsc, select= -c(POC, DSC_website, source, cruise_report_site_char))

#Include only proejcts from the PAC (82 projects)
dsc2<-dsc1[dsc1$PAC == '1', ]

#Replace ?, ??, ?? with NAs
dsc2[ dsc2 == c("?", "??", "???") ] <- NA

# Get rid of dollar signs
dsc2$cost = as.numeric(gsub("\\$", "", dsc2$cost))

#Replace NA values with 0
dsc2$cost[is.na(dsc2$cost)] <- 0
```


Troubles with the str function
```{r}
str(dsc2)
```


##Plotting geospatial data
*This could be used to visualize information about fishery regions: amount funding, objectives, NOMEC categories, ect.*
https://www.paulamoraga.com/book-geospatial/sec-flexdashboard.html 

#Set up 
install.packages("rnaturalearth")
devtools::install_github("ropensci/rnaturalearthhires")
install.packages("wbstats")
```{r}
#Create map object which includes countries name and iso_a3
library(rnaturalearth)
library(rnaturalearthhires)
map <- ne_download() #this is the only part that actually is the map 
names(map)[names(map) == "iso_a3"] <- "ISO3"
names(map)[names(map) == "name"] <- "NAME"

#plot the map object
library(sp)
plot(map)



#This step is retrieving data from World Bank - I will not need to do this step for my version
library(wbstats)
indicators <- wbsearch(pattern = "pollution")


#Select indicators to plot- for me I may chose funding, and maybe project objectives? Could have to reformat
d <- wb(
  indicator = "EN.ATM.PM25.MC.M3", #decide to plot the indicators air pollution and mean annual exposure 
  startdate = 2016, enddate = 2016 #for the year 2016
)

map$PM2.5 <- d[match(map$ISO3, d$iso3c), "value"] #match country codes to PM2.5 calues



```

#Create dashboard table
install.packages("DT")
```{r}
#Create the intereactive table
library(DT)
DT::datatable(map@data[, c("ISO3", "NAME", "PM2.5")], #create a data table showing the following 3 variables
  rownames = FALSE, options = list(pageLength = 10) #don't show row names  and page length is set to 10 rows
)
```

#Create interactive map 
install.packages("leaflet")
```{r}
library(leaflet)

pal <- colorBin(
  palette = "viridis", domain = map$PM2.5, #color the countries according to their PM2.5 value- change domain for yours
  bins = seq(0, max(map$PM2.5, na.rm = TRUE) + 10, by = 10) #sequence ranges from 0 to max PM values in increments of 10
)


map$labels <- paste0(
  "<strong> Country: </strong> ",
  map$NAME, "<br/> ",
  "<strong> PM2.5: </strong> ",
  map$PM2.5, "<br/> "
) %>%
  lapply(htmltools::HTML) #write labels using html codes to show PM levels of each country- apply then to argument add polygon

leaflet(map) %>%
  addTiles() %>% #adds background map
  setView(lng = 0, lat = 30, zoom = 2) %>% #center and zoom map- I can use this to center the map over the US
  addPolygons(
    fillColor = ~ pal(PM2.5), #color the areas by given PM2.5 values set in "pal" object
    color = "white", #color border with white to see background
    fillOpacity = 0.7,#thickness of border
    label = ~labels, #add labels- code written in html above  
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  ) %>%
  leaflet::addLegend( #add legend
    pal = pal, values = ~PM2.5,
    opacity = 0.7, title = "PM2.5"
  )

```

Notes: 
- I need to create a column for this that indicates the region in dataset
- I need to decide if I want to color a fisheries region, or the states associated with the fisheries region... I feel like the fisheries region itself would make more sense
- explore what it would look like if I had the map being colored for different things: cost, objectives, NOMEC, ect.
*ask Heather

##Creating a pie chart
*This could be used to show the proportion of different obectives funded over the years/ could compliment map*
https://stackoverflow.com/questions/50319450/how-to-create-a-pie-chart-with-rstudio-flexdashboard 

I like the idea of using a pie chart to show proportions of objectives funded over the years.

Notes: 
-I will have to group projects by start funding year
- then pivot longer the objectives columns into one row with counts for the year
- I want the variable that can be changed ontop to be year 

```{r}
library("plotly")

totalref <- read.table(header = T, text = 
"Country        Refugees   
'Belgium'         508645
'France'         2874490
'Germany'       10893516
'Ireland'         142555
'Italy'           538407
'Netherlands'    1900518
'Portugal'          7918
'Spain'            89946
'United Kingdom' 3766855")

plot_ly(totalref, labels = ~ Country, values = ~ Refugees, type = 'pie',
        textposition = 'inside', textinfo = 'label+percent') %>%
    layout (title='Persons of Concern in 9 European Countries',
            showlegend = TRUE)
```

```{r}
data_offices<- dsc2[,8:32]
colSums(data_offices)
data_offices
table(dsc2$NMFS)#52
table(dsc2$SEFSC) #3
table(dsc2$NEFSC) #3
table(dsc2$SWFSC) #35
table(dsc2$NWFSC) #30
table(dsc2$PIFSC) #1
table(dsc2$AFSC) #4
table(dsc2$OST) #3
table(dsc2$`NSL/SI`) #1
table(dsc2$`Council(s)`) #3
table(dsc2$NOS) #41
table(dsc2$NCCOS) #23
table(dsc2$ONMS) #35
table(dsc2$FGBNMS) #3
table(dsc2$GRNMS) #3
table(dsc2$CINMS) #25
table(dsc2$MBNMS) #4
table(dsc2$CBNMS) #9
table(dsc2$GFNMS) #6
table(dsc2$OCNMS) #4
table(dsc2$PMNM) #3
table(dsc2$OAR) #4
table(dsc2$OER) #3
table(dsc2$NESDIS) #1
table(dsc2$NCEI) #1


library("plotly")


office<- read.table(header = T, text = 
"NOAA Office        Funded   
'NMFS'               51
'SEFSC'              3
'NEFSC'              3
'SWFSC'              35
'NWFSC'              30
'PIFSC'              1
'AFSC'               4
'OST'                3
   

")

plot_ly(totalref, labels = ~ Country, values = ~ Refugees, type = 'pie',
        textposition = 'inside', textinfo = 'label+percent') %>%
    layout (title='Persons of Concern in 9 European Countries',
            showlegend = TRUE)

library(plyr)
library(data.table)
count(dsc2, 'NMFS')

dsc2[, sum(freq), by=c(NMFS, NCEI)]
View(dsc2)

colSums(dsc2[ , c("NMFS", "NCEI", "NESDIS")])
```
```

#Bar graph
*The accompanying bargraph could show which NOAA offices were funded? This might not work well because there are so many different ones? Maybe if I oriented it sideways it wouldn't be so crowded?.. that is 27 variables*

- maybe look at choosing line offices 
- NCCOS, OAR, ect. pick fewer than 10 that are the most interesting 
- make NMOS, NCCOS, and science centers should be on there 

- Is there a way that we could group all of these into smaller divisions? Fisheries, oceans, ect.?

# Playing around with Functions

*Structure of a function* 
my.func <- function(arg1, arg2, arg3, ...) {
<commands>
return(output.object)
}

The code below is an example of returning a single variable in a function :

Translated, this function says “if you give me an object, that I
will call x, I will store its mean() as meanval, then its sd() as sdval,
and then return their ratio sdval/meanval.”
```{r}
coef.of.var <- function(x){
meanval <- mean(x,na.rm=TRUE) # recall this means "ignore NAs"
sdval <- sd(x,na.rm=TRUE)
return(sdval/meanval)
}
```

Now use the function with the 1973 New York ozone data 
```{r}
data(airquality) # make the data available in this R session
coef.of.var(airquality$Ozone)
```

*Example: Returning multiple values

A function can *return multiple objects/values by using list()* – which collects objects of (potentially) different types.

The function below calculates estimates of the mean and standard deviation of a population, based on a vector (x) of observations;

```{r}
popn.mean.sd <- function(x){
n <- length(x)
mean.est <- mean(x,na.rm=TRUE)
var.est <- var(x,na.rm=TRUE)*(n-1)/n
est <- list(mean=mean.est, sd=sqrt(var.est))
return(est)
}
```

Applying our popn.mean.sd() function to the daily ozone concentrations in New York data;
```{r}
results <- popn.mean.sd(airquality$Ozone)
 attributes(results) #list the attributes of the object returned

results$mean
results$sd


#With my own data 

results1 <- popn.mean.sd(dsc2$cost)
 attributes(results1) 
 
 results1$mean
 

dsc2$project_ID<- as.character(dsc2$project_ID)
dsc2$NMFS<- as.numeric(dsc2$NMFS)
str(dsc2)
 
```
 
*Declaring functions with in functions*
Usually, functions that take arguments, execute R commands, and return output will be enough. But functions can be declared and used inside a function;

Translated; “if you given me a number, that I will call y, I will
define a function I call square that takes a number that it calls x
and returns x-squared, then similarly one I call cube that cubes,
then I will return the sum of applying square to y and cube to y”.

```{r}
square.plus.cube <- function(y) {
square <- function(x) { return(x*x) }
cube <- function(x) { return(x^3) }
return(square(y) + cube(y))
}

#Example
square.plus.cube(4) #answer is 80
```


Using plotly() to create a map: 
https://cran.r-project.org/web/packages/usmap/vignettes/mapping.html 
```{r}
library(usmap)
library(ggplot2)

plot_usmap(regions = "counties") + 
  labs(title = "US Counties",
  
       subtitle = "This is a blank map of the counties of the United States.") + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue"))
```


 
Note that the `echo = FALSE` parameter was added to the code chunk to prevent
printing of the R code that generated the plot.
