---
title: "Shiny Tutorials"
author: "Sydney Lewis"
date: "2022-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Started Tutorial 
http://rstudio.github.io/shinydashboard/get_started.html


## Installation 

install.packages("shinydashboard")
devtools::install_github("shiny", "rstudio")

## Basics

A dashboard has three parts: a header, a sidebar, and a body. Here’s the most minimal possible UI for a dashboard page.

```{r eval=FALSE, include=FALSE}
## ui.R ##
library(shiny)
library(shinydashboard)

dashboardPage(
  dashboardHeader(), #header- where you can name your dashboard
  dashboardSidebar(), #sidebar- where you can make tabs
  dashboardBody() #body- where graphs go 
)
```

## Blank dashboard

Obviously, this dashboard isn’t very useful. We’ll need to add components that actually do something. In the body we can add boxes that have content.

```{r eval=FALSE, include=FALSE}
## app.R ##
library(shiny)
library(shinydashboard)

ui <- dashboardPage(
  dashboardHeader(title = "Basic dashboard"), #"Deep-Sea Coral dashboard"
  dashboardSidebar(), 
  dashboardBody( #this sets the stage for inserting plots and such into the dashboard 
    # Boxes need to be put in a row (or column)
     fluidRow(
      box(plotOutput("plot1", height = 250)), #I am not sure what plot1  

      box(
        title = "Controls",
        sliderInput("slider", "Number of observations:", 1, 100, 50)
      )
    )
  )
)

server <- function(input, output) {
  set.seed(122)
  histdata <- rnorm(500)

  output$plot1 <- renderPlot({
    data <- histdata[seq_len(input$slider)]
    hist(data)
  })
}

#shinyApp(ui, server)
```

## Basic dashboard

Next, we can add content to the sidebar. For this example we’ll add menu items that behave like tabs. These function similarly to Shiny’s tabPanels: when you click on one menu item, it shows a different set of content in the main body.

There are two parts that need to be done. First, you need to add menuItems to the sidebar, with appropriate tabNames.

```{r}
ui2<-
  dashboardPage(
  dashboardHeader(title = "Basic dashboard"), #create the title of the dashboard
  dashboardSidebar( #labels for side tabs
    sidebarMenu(
      menuItem("Dashboard", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Widgets", tabName = "widgets", icon = icon("th"))
    )
  ),
  dashboardBody(
    tabItems( #list of visuals that will be included when you click on each new tab
      # First tab content
      tabItem(tabName = "dashboard", #correspond tabName to sidebarMenu
        fluidRow( 
          box(plotOutput("plot1", height = 250)), #dimensions of graph
          box(
            title = "Controls", #title of slidebar 
            sliderInput("slider", "Number of observations:", 1, 100, 50)
          )
        )
      ),
  # Second tab content
      tabItem(tabName = "widgets", 
        h2("Widgets tab content") #the words displayed on the Widgets tab
      )
)
)
)

##This code is needed to help render the dashboard outside of R 
server <- function(input, output) {
  set.seed(122)
  histdata <- rnorm(500)

  output$plot1 <- renderPlot({
    data <- histdata[seq_len(input$slider)]
    hist(data)
  })
}

shinyApp(ui2, server)

```

#Background: Shiny and HTML

```{r}
# A basic div
div(class = "my-class", "Div content")
```

Div content
```{r}
# Nested HTML tags
div(class = "my-class", p("Paragraph text"))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
